Mass and occupation fraction of dark matter halos hosting Lyman-$\alpha$ emitters at $z\sim 3$

### 2013-12-31 paper/ms.tex
fixed typos

\title[Halo mass and occupation fraction of LAEs at$z=3.1$]{Mass and
  occupation fraction of dark matter halos hosting Lyman-$\alpha$
  emitters at $z\sim 3$}      
\author[~J.~E. Forero-Romero and ~J.~E. Mejia-Restreo]{
\parbox[t]{\textwidth}{\raggedright 
  Jaime E. Forero-Romero$^{1}$ and
  Julian E. Mej\'ia-Restrepo$^{2}$ 
}
\vspace*{6pt}\\
$^{1}$ Departamento de F\'{i}sica, Universidad de los Andes, Cra. 1
No. 18A-10, Edificio Ip, Bogot\'a, Colombia \\
$^{2}$ Departamento de Astronom\'{i}a, Universidad de Chile, Camino el
Observatorio 1515, Santiago, Chile}

@@

\begin{abstract}
%
We derive constraints on the mass and occupation fraction of dark
matter halos hosting \ly Emitting galaxies (LAEs) at a redshift of
$z=3.1$ by matching the number density and the angular
correlation function between mock and observed fields. We explicitly
take into account the cosmic variance on the typical observed field
size by constructing mock fields from a large cosmological N-body
simulation matching the geometries of observed fields. To populate the
halos in the simulation we use a model where a dark matter halo with
mass in the range $M_{\rm min}<M_{\rm h}<M_{\rm max}$ can only host
one detectable LAE with a probability $f_{\rm occ}$. The main motivation to
build such a model is to avoid the large uncertainties in the
estimation of \ly luminosities. Our main result is that the clustering
and number density information are insufficient to impose a tight
constraint on the {\bf occupation} fraction. On the other hand the
minimum mass and maximum mass can be tightly constrained to the range
$M_{\rm max}<10^{12}$\hMsun and $10^{10}\hMsun\leq M_{\rm min}\leq
10^{11.5}\hMsun$.  We also find that the consistent models have a narrow mass
range $\Delta M \equiv \log_{10}M_{\rm max} - \log_{10} M_{\rm min}$
smaller than $1.0$ dex, with the vast majority in the range $\Delta
M<0.5$ dex. Already published constraints on the mass range and
occupation fraction are a sub-set of the models we report in this
\documentname. Our results present an inconclusive response as to
which is the occupation fraction for LAEs at $z=3.1$. A precise answer
requires observations on larger fields and robust physical models for
the connection between dark matter, star formation and the escape of
\ly radiation. That kind of modelling should also be able to provide a
plausible explanation for the narrow range in mass $\Delta M<1.0$ dex
derived here for the halos hosting LAEs at $z=3.1$.

@@

Capitalizing LAEs observations requires an understanding of
their place in the context of a given structure formarion model in a
cosmological context. Under the current paradigm the dominant matter
content of the Universe is to be found in dark matter (DM) and each
galaxy is thought to be hosted by larger dark matter structure known
as a halo. \citep{Peebles1980,SpringelNature05}.

@@

{\bf A secondary benchmark is the angular correlation function
(ACF). \cite{Yamada2012} does not report an ACF measurement for
  their fields. Instead we use the results by   \cite{Ouchi2008} who
  reported the ACF is over a region of $1$ deg$^2$ over the SXDS
  field.  We also keep track of the results by \cite{Hayashino2004}
  who reported the ACF in the SSA22 region, including its densest
  subfield.}

@@

{\bf The field SSA22 has been known to harbor a significant
  galaxy overdensity. \cite{Yamada2012} estimates that the densest
  subfield is likely to be a rare density peak with $3-4\sigma$
  significance on the scale of $\sim 60$\hMpc. }

{\bf All the other sub-fields in SSA22 are average. They have a similar
  number density as other blank fields. The left panel in Figure
  \ref{fig:halos} shows the integrated distribution of LAEs number
  density in all the fields reported in  \cite{Yamada2012}. It is
  evident that there is only one sub-field  that stands out as an
  outlier in the distribution (it corresponds to  the densest field in
  SSA22). All the other fields form a continous  distribution in
  number density.}

@@

This plot allows us to understand why only a specific range of
models ${\mathcal M}$ can be expected to be consistent with
observations. From Figure \ref{fig:halos} we can read that models with
a minimum mass $M_{\rm min}>10^{12}$\hMsun always have a
surface number density lower than the observational constrain, making
them incompatible with observations; there are simply to few halos compared to observed
LAEs. The opposite is true in models with $M_{\rm
  min}<10^{11.0}$\hMsun that have a surface number density larger
observations. In those cases the maximum mass $M_{\rm max}$ and the
occupation fraction $f_{\rm occ}<1.0$  can be tuned in order to lower
the halo number density to match observations.

@@

Figure \ref{fig:correlation_parameters} shows the results for the
best estimates of $\theta_{0}$-$\beta$  used in the ACF
parameterization.  Blue circles with error bars represent the results
from the mocks and the horizontal line the obsrvational results of
\cite{Ouchi2010} over a field with average number density where the
$\beta$ parameter was fixed in the fit. From this Figure we observe
that there are models with too large values of $\theta_0$ and $\beta$
that can be ruled out as inconsistent with the angular correlation
function.

@@

\begin{figure*}
\begin{center}
\includegraphics[width=0.46\linewidth,angle=0]{Fig6_mass.pdf}
\hspace{5mm}
\includegraphics[width=0.46\linewidth,angle=0]{Fig6_f_occ.pdf}
\end{center}
\caption{Planes $M_{\rm min}$-$M_{\rm max}$ (left) and $M_{\rm
    min}$-$f_{\rm occ}$ (right). with the models fulfilling both
   constraints on the maximal number of consistent mocks and the
  angular correlation function. The observational data correspods to
  \citet{Ouchi2010}.   
  \label{fig:restriction_mock_and_f_occ_corr}} 
\end{figure*}

Out of the initial set of $9000$ models we end up with $40$ that are
consistent with the observational constraints. To facilitate the
discussion of these models we define a new quantity, the halo mass
range $\Delta M=\log_{10}M_{\rm max} - \log_{10}M_{\rm  min}$, which
together with the occupation fraction, $f_{\rm occ}$, and the minimum
mass $M_{\rm min}$ allows us to clasify all the successful models into
three families:

@@

{\bf The models in the third family are barely consistent with the
  constraints from the ACF. They have mean values of $\theta_0$
  outside the observational uncertainties, but have
  are considered as consistent because their large error barsoverlap. These
  models they have another particular feature: their minimum mass is
  exactly $M_{\rm min}=10^{10.9}\hMsun$. Because these models are
  are barely compatible with observations and are a minority of the
  consistent models, we exclude them from the rest of the discussion.}

@@

In summary, we show that {\bf the currently available spatial information of LAEs cannot
put a tight constraint on their occupation fraction of dark matter
halos}. This information can confidently be used to stablish that
these galaxies are hosted in low mass halos with masses within a range
spanning less than a factor of ten. We foresee that the new
observations with new instruments (such as MUSE, Hyper SuprimeCam and
HETDEX) covering larger fields and a wider range of luminosities will
be key in reducing cosmic variance and imposing tighter constraints on
the properties of dark matter halos hosting LAEs. At the same time,
additional modeling for \ly radiation transfer is needed to put a
tighter constrain on the \ly escape fraction in high redshift
galaxies, also paying attention to newly framed physical phenomena,
such as the stochasticity \citep{ForeroRomero2013} in the star
formation process, which might play a role in inducing detection
biases in high redshift LAEs.

>>>>

\title[Halo mass and occupation fraction of LAEs at$z=3.1$]{Mass and
  occupation fraction of dark matter halos hosting Lyman-$\alpha$
  emitters at $z\sim 3$}      
\author[~J.~E. Forero-Romero and ~J.~E. Mejia-Restrepo]{
\parbox[t]{\textwidth}{\raggedright 
  Jaime E. Forero-Romero$^{1}$ and
  Julian E. Mej\'ia-Restrepo$^{2}$ 
}
\vspace*{6pt}\\
$^{1}$ Departamento de F\'{i}sica, Universidad de los Andes, Cra. 1
No. 18A-10, Edificio Ip, Bogot\'a, Colombia \\
$^{2}$ Departamento de Astronom\'{i}a, Universidad de Chile, Camino el
Observatorio 1515, Santiago, Chile}

@@

\begin{abstract}
%
We derive constraints on the mass and occupation fraction of dark
matter halos hosting \ly Emitting galaxies (LAEs) at a redshift of
$z=3.1$ by matching the number density and the angular
correlation function between mock and observed fields. We explicitly
take into account the cosmic variance on the typical observed field
size by constructing mock fields from a large cosmological N-body
simulation matching the geometries of observed fields. To populate the
halos in the simulation we use a model where a dark matter halo with
mass in the range $M_{\rm min}<M_{\rm h}<M_{\rm max}$ can only host
one detectable LAE with a probability $f_{\rm occ}$. The main motivation to
build such a model is to avoid the large uncertainties in the
estimation of \ly luminosities. Our main result is that the clustering
and number density information are insufficient to impose a tight
constraint on the {\bf occupation} fraction. On the other hand the
minimum mass and maximum mass can be tightly constrained to the range
$M_{\rm max}<10^{12}$\hMsun and $10^{10}\hMsun\leq M_{\rm min}\leq
10^{11.5}\hMsun$.  We also find that the consistent models have a narrow mass
range $\Delta M \equiv \log_{10}M_{\rm max} - \log_{10} M_{\rm min}$
smaller than $1.0$ dex, with the vast majority in the range $\Delta
M<0.5$ dex. Already published constraints on the mass range and
occupation fraction are a sub-set of the models we report in this
\documentname. Our results present an inconclusive response as to
which is the occupation fraction for LAEs at $z=3.1$. A precise answer
requires observations on larger fields and robust physical models for
the connection between dark matter, star formation and the escape of
\ly radiation. That kind of modeling should also be able to provide a
plausible explanation for the narrow range in mass $\Delta M<1.0$ dex
derived here for the halos hosting LAEs at $z=3.1$.

@@

Capitalizing LAEs observations requires an understanding of
their place in the context of a given structure formation model in a
cosmological context. Under the current paradigm the dominant matter
content of the Universe is to be found in dark matter (DM) and each
galaxy is thought to be hosted by larger dark matter structure known
as a halo. \citep{Peebles1980,SpringelNature05}.

@@

{\bf A secondary benchmark is the angular correlation function
(ACF). \cite{Yamada2012} does not report an ACF measurement for
  their fields. Instead we use the results by   \cite{Ouchi2008} who
  reported the ACF is over a region of $1$ deg$^2$ over the SXDS
  field.  We also keep track of the results by \cite{Hayashino2004}
  who reported the ACF in the SSA22 region, including its densest
  sub-field.}

@@

{\bf The field SSA22 has been known to harbor a significant
  galaxy overdensity. \cite{Yamada2012} estimates that the densest
  sub-field is likely to be a rare density peak with $3-4\sigma$
  significance on the scale of $\sim 60$\hMpc. }

{\bf All the other sub-fields in SSA22 are average. They have a similar
  number density as other blank fields. The left panel in Figure
  \ref{fig:halos} shows the integrated distribution of LAEs number
  density in all the fields reported in  \cite{Yamada2012}. It is
  evident that there is only one sub-field  that stands out as an
  outlier in the distribution (it corresponds to  the densest field in
  SSA22). All the other fields form a continuous  distribution in
  number density.}

@@

This plot allows us to understand why only a specific range of
models ${\mathcal M}$ can be expected to be consistent with
observations. From Figure \ref{fig:halos} we can read that models with
a minimum mass $M_{\rm min}>10^{12}$\hMsun always have a
surface number density lower than the observational constrain, making
them incompatible with observations; there are simply too few halos compared to observed
LAEs. The opposite is true in models with $M_{\rm
  min}<10^{11.0}$\hMsun that have a surface number density larger
observations. In those cases the maximum mass $M_{\rm max}$ and the
occupation fraction $f_{\rm occ}<1.0$  can be tuned in order to lower
the halo number density to match observations.

@@

Figure \ref{fig:correlation_parameters} shows the results for the
best estimates of $\theta_{0}$-$\beta$  used in the ACF
parameterization.  Blue circles with error bars represent the results
from the mocks and the horizontal line the observational results of
\cite{Ouchi2010} over a field with average number density where the
$\beta$ parameter was fixed in the fit. From this Figure we observe
that there are models with too large values of $\theta_0$ and $\beta$
that can be ruled out as inconsistent with the angular correlation
function.

@@

\begin{figure*}
\begin{center}
\includegraphics[width=0.46\linewidth,angle=0]{Fig6_mass.pdf}
\hspace{5mm}
\includegraphics[width=0.46\linewidth,angle=0]{Fig6_f_occ.pdf}
\end{center}
\caption{Planes $M_{\rm min}$-$M_{\rm max}$ (left) and $M_{\rm
    min}$-$f_{\rm occ}$ (right). with the models fulfilling both
   constraints on the maximal number of consistent mocks and the
  angular correlation function. The observational data corresponds to
  \citet{Ouchi2010}.   
  \label{fig:restriction_mock_and_f_occ_corr}} 
\end{figure*}

Out of the initial set of $9000$ models we end up with $40$ that are
consistent with the observational constraints. To facilitate the
discussion of these models we define a new quantity, the halo mass
range $\Delta M=\log_{10}M_{\rm max} - \log_{10}M_{\rm  min}$, which
together with the occupation fraction, $f_{\rm occ}$, and the minimum
mass $M_{\rm min}$ allows us to classify all the successful models into
three families:

@@

{\bf The models in the third family are barely consistent with the
  constraints from the ACF. They have mean values of $\theta_0$
  outside the observational uncertainties, but have
  are considered as consistent because their large error bars overlap. These
  models they have another particular feature: their minimum mass is
  exactly $M_{\rm min}=10^{10.9}\hMsun$. Because these models are
  are barely compatible with observations and are a minority of the
  consistent models, we exclude them from the rest of the discussion.}

@@

In summary, we show that {\bf the currently available spatial information of LAEs cannot
put a tight constraint on their occupation fraction of dark matter
halos}. This information can confidently be used to establish that
these galaxies are hosted in low mass halos with masses within a range
spanning less than a factor of ten. We foresee that the new
observations with new instruments (such as MUSE, Hyper SuprimeCam and
HETDEX) covering larger fields and a wider range of luminosities will
be key in reducing cosmic variance and imposing tighter constraints on
the properties of dark matter halos hosting LAEs. At the same time,
additional modeling for \ly radiation transfer is needed to put a
tighter constrain on the \ly escape fraction in high redshift
galaxies, also paying attention to newly framed physical phenomena,
such as the stochasticity \citep{ForeroRomero2013} in the star
formation process, which might play a role in inducing detection
biases in high redshift LAEs.
